The hydra and the rooster

### 2020-07-22 latex/chapter-hydras.tex
renaming fight --> battle

For this purpose, we define two mutual inductive types.
\index{Types!Hydra}

@@

@@

In this section, we represent the rules of hydra battles as a binary relation associated with
a \emph{round}, i.e., an interaction composed of the two following actions:
\begin{enumerate}
\item Hercules chops off one head of the hydra
\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
\end{enumerate}
The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot).

@@

Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into several cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.

We decompose the relation associated with each  round in two cases: 
\begin{itemize}
\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
\end{itemize}

@@

\emph{Please consider exercise~\vref{exo:interactive-battle}.}

@@

We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$.

@@

\subsection{Fights}

@@

\begin{Coqsrc}
Inductive fight (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
|  fight_1 : forall i h  h', battle_r   B i  h h' -> 
                            fight B i h (S i) h'
| fight_n : forall i h  j h' h'',  battle_r  B i h h''  ->
                                   fight B (S i) h'' j h'  ->
                                   fight B i h j h'.
\end{Coqsrc}

@@

%  \begin{Coqsrc}
%  Lemma fight_free_equiv1 : forall i j h h',  
%              fight free i h j h' ->   h -+-> h'.

%  Lemma fight_free_equiv2 : forall h h',
%      h -+-> h' ->
%     forall i, exists j,  fight free i h j h'.
%  \end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_0_2 : fight standard 0 hinit 2 (hyd1 h3).   
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_2_3 : fight standard 2 (hyd1 h3)  3 (hyd 3 0 0).

Lemma L_0_3 : fight standard 0 hinit 3 (hyd 3 0 0).
\end{Coqsrc}

@@

The following function returns the next configurarion of the game. 
Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
\texttt{lemma:step-fight}, page~\pageref{lemma:step-fight}.

@@

\label{lemma:step-fight}

\begin{Coqsrc}
Lemma step_fight : forall i a b c a' b' c', 
   one_step i a b c a' b' c' ->
   fight standard i (hyd  a b c)  (S i) (hyd a' b' c').
\end{Coqsrc}

@@

The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{fight}.

\begin{Coqsrc}
 Lemma steps_fight : forall i a b c j a' b' c', 
   steps i a b c j a' b' c' ->
   fight standard i (hyd  a b c)   j  (hyd a' b' c').
\end{Coqsrc}

Thus, any result about \texttt{steps} will be applicable to standard fights.
Using the predicate \texttt{steps} our study of the length of the considered battle
can  be decomposed into three parts:

@@

From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
the fight.

@@

We are now able to prove formally that the considered fight is 
composed of $N$ steps.

\begin{Coqsrc}
Lemma Almost_done :
  fight standard 3 (hyd 3 0 0) N (hyd 0 0 0).
Proof. 
  apply steps_fight, SuperbigStep.
Qed.

Theorem  Done :
  fight standard 0 hinit N head.
Proof.
  eapply fight_trans.
  -   apply Almost_done.
  -  apply L_0_3.
Qed.
\end{Coqsrc}

>>>>

For this purpose, we define two mutual \emph{ad-hoc}  inductive types, where \texttt{Hydra} is the main type, and \texttt{Hydrae} is a helper for describing finite sequences of hydra.
\index{Types!Hydra}

@@

\index{To do}

\begin{todo}
Look for an existing library on trees with nodes of arbitrary arity, iin order to replace  this ad-hoc type with something more generic.
\end{todo}

@@

In this section, we represent the rules of hydra battles as a binary relation associated with
a \emph{round}, i.e., an interaction composed of the two following actions:
\begin{enumerate}
\item Hercules chops off one head of the hydra
\item Then, the  hydra replicates the wounded part (if the head is at distance $\geq 2$ from the foot).
\end{enumerate}
The relation associated with each round of the battle is \emph{parameterized}  by the \emph{expected} replication  factor (irrelevant if the chopped head is at distance 1 from the foot,
but present for consistency's sake).

@@

Since the replication of parts of the hydra depends on the distance of the chopped head from  the foot, we  decompose our description into two main  cases, under the form of a bunch of [mutually] inductive predicates over the types \texttt{Hydra} and \texttt{Hydrae}.

The mutually exclusive cases we consider are the following.
\begin{itemize}
\item \textbf{R1}: The chopped off head was at distance 1 from the foot.
\item \textbf{R2}: The chopped off head was at a distance greater than or equal to  $2$ from the foot.
\end{itemize}

@@

The reader is encouraged to look at all the intermediary subgoals of this example.
\emph{Please consider also exercise~\vref{exo:interactive-battle}.}

@@

We chosed to call \emph{standard} the kind of battles which appear  most often in the litterature and correspond to an arithmetic progression of the replication factor : $0,1,2,3, \dots$

@@

\subsection{Big steps}

@@

\begin{Coqsrc}
Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
|  battle_1 : forall i h  h', battle_r   B i  h h' -> 
                            battle B i h (S i) h'
| battle_n : forall i h  j h' h'',  battle_r  B i h h''  ->
                                   battle B (S i) h'' j h'  ->
                                   battle B i h j h'.
\end{Coqsrc}

@@

%  \begin{Coqsrc}
%  Lemma battle_free_equiv1 : forall i j h h',  
%              battle free i h j h' ->   h -+-> h'.

%  Lemma battle_free_equiv2 : forall h h',
%      h -+-> h' ->
%     forall i, exists j,  battle free i h j h'.
%  \end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).

Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
\end{Coqsrc}

@@

The following function returns the next configurarion of the game. 
Note that this function is defined only for making experiments and is not  ``certified''.  Formal proofs about our battle will only start with the lemma
\texttt{lemma:step-battle}, page~\pageref{lemma:step-battle}.

@@

\label{lemma:step-battle}

\begin{Coqsrc}
Lemma step_battle : forall i a b c a' b' c', 
   one_step i a b c a' b' c' ->
   battle standard i (hyd  a b c)  (S i) (hyd a' b' c').
\end{Coqsrc}

@@

The following lemma establishes a relation between \texttt{steps} and the predicate \texttt{battle}.

\begin{Coqsrc}
 Lemma steps_battle : forall i a b c j a' b' c', 
   steps i a b c j a' b' c' ->
   battle standard i (hyd  a b c)   j  (hyd a' b' c').
\end{Coqsrc}

Thus, any result about \texttt{steps} will be applicable to standard battles.
Using the predicate \texttt{steps} our study of the length of the considered battle
can  be decomposed into three parts:

@@

From now on, the lemma \texttt{reachable\_S} allows us to watch larger steps of 
the battle.

@@

We are now able to prove formally that the considered battle is 
composed of $N$ steps.

\begin{Coqsrc}
Lemma Almost_done :
  battle standard 3 (hyd 3 0 0) N (hyd 0 0 0).
Proof. 
  apply steps_battle, SuperbigStep.
Qed.

Theorem  Done :
  battle standard 0 hinit N head.
Proof.
  eapply battle_trans.
  -   apply Almost_done.
  -  apply L_0_3.
Qed.
\end{Coqsrc}

### 2020-07-22 latex/chapter-ks.tex
renaming fight --> battle

\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1}$ in one round with
the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) fight from $\iota(\alpha)$ to $\iota(\beta)$.
\end{itemize}

@@

The next step of our development is to extend this relationship to
the order $<$ on $[0,\epsilon_0[$ on one side, and hydra fights on the other side.

\begin{Coqsrc}
Lemma path_to_fight alpha s beta :
  path_to  beta  s alpha -> nf alpha ->
  iota alpha -+-> iota beta.
\end{Coqsrc}

As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) fight.

\begin{Coqsrc}
Lemma LT_to_fight alpha beta :
    beta < alpha ->  iota alpha -+-> iota beta.
\end{Coqsrc}

@@

\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h < mu}
\item By Lemma \texttt{LT\_to\_fight}, we get a (free) battle from
\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.

@@

\begin{Coqsrc}
Lemma LT_to_standard_battle :
    forall alpha beta,
      beta < alpha ->
      exists n i,  fight standard  n (iota alpha) i (iota beta).
\end{Coqsrc}

>>>>

\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1}$ in one round with
the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
\end{itemize}

@@

The next step of our development is to extend this relationship to
the order $<$ on $[0,\epsilon_0[$ on one side, and hydra battles on the other side.

\begin{Coqsrc}
Lemma path_to_battle alpha s beta :
  path_to  beta  s alpha -> nf alpha ->
  iota alpha -+-> iota beta.
\end{Coqsrc}

As a corollary, we are now able to transform any inequality $\beta<\alpha<\epsilon_0$ into a (free) battle.

\begin{Coqsrc}
Lemma LT_to_battle alpha beta :
    beta < alpha ->  iota alpha -+-> iota beta.
\end{Coqsrc}

@@

\item From the hypothesis \texttt{m\_bounded}, we have \texttt{m big\_h < mu}
\item By Lemma \texttt{LT\_to\_battle}, we get a (free) battle from
\texttt{big\_h = iota mu} to \texttt{small\_h = iota (m big\_h)}.

@@

\begin{Coqsrc}
Lemma LT_to_standard_battle :
    forall alpha beta,
      beta < alpha ->
      exists n i,  battle standard  n (iota alpha) i (iota beta).
\end{Coqsrc}

### 2020-07-22 latex/chapter-large-sets.tex
renaming fight --> battle

In this chapter, we try to feel how long a standard battle can be.
To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
we give a minoration of the number of steps of a standard fight which
starts with the hydra $\iota(\alpha)$ and the replication factor $k$.

@@

\begin{Coqsrc}
Theorem battle_length_std (alpha : E0)  :
  alpha <> Zero ->
  forall k, (1 <= k)%nat ->
            fight_length standard k (iota (cnf alpha))
                         (L_ alpha (S k) - k).
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H_ alpha k - k <= l /\
       fight_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

>>>>

In this chapter, we try to feel how long a standard battle can be.
To be precise, for any ordinal $\alpha<\epsilon0$ and any positive integer $k$,
we give a minoration of the number of steps of a standard battle which
starts with the hydra $\iota(\alpha)$ and the replication factor $k$.

@@

\begin{Coqsrc}
Theorem battle_length_std (alpha : E0)  :
  alpha <> Zero ->
  forall k, (1 <= k)%nat ->
            battle_length standard k (iota (cnf alpha))
                         (L_ alpha (S k) - k).
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H_ alpha k - k <= l /\
       battle_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

### 2020-07-22 latex/chapter-ordinals.tex
renaming fight --> battle

For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
build a fight transforming \texttt{big\_h} into \texttt{small\_h}.

@@

The following lemma establishes a correspondance between the relation
\texttt{step} and hydra fights.

\begin{Coqsrc}
Lemma step_to_fight : forall p q, step p q -> iota p -+-> iota q.
\end{Coqsrc}

@@

Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
by transfinite induction over \texttt{p} a fight 
that transforms the hydra $\iota(p)$ into $\iota(q)$.

>>>>

For proving the inequality  \texttt{m\_lt: m small\_h < m big\_h}, it suffices to
build a battle transforming \texttt{big\_h} into \texttt{small\_h}.

@@

The following lemma establishes a correspondance between the relation
\texttt{step} and hydra battles.

\begin{Coqsrc}
Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
\end{Coqsrc}

@@

Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
by transfinite induction over \texttt{p} a battle 
that transforms the hydra $\iota(p)$ into $\iota(q)$.
